\section{Abstract}
Through the rise of so called \emph{Big Data} as well as the corresponding analytics, the requirements for fast and efficient data retrieval and manipulation are suggesting new advancement with parsers. Those conditions make parsing techniques regain importance in order to allow faster data access and treatment. However constructing parse trees can be expensive as not all of the parsed content might be of interest or if the data in its basic form is not useful and needs to be transformed anyway.
We will present a generic parsing framework in Scala which can be used to compose a formal grammar with different semantic functions who are applied directly during the parsing process. The \textbf{AGParsers} (attribute grammar parsers) use a polyvalent structure to allow to user to easily extend it and adapt it to his own use-cases.
We use ideas from attribute grammars to combine syntactic and semantic features during parsing thus creating a parsing framework which can carry partial result information through the parse tree which can be used directly during parsing. Thus the user can define exactly which structure he would like to be created by the parser without the need of several manipulating passes over the parse tree structure.
The described functionality is achieved by augmenting the existing scala parser combinator to be able to carry additional information during parsing which can be used and augmented depending on each parsing rule and the related semantic attribute function.
